Old is Still Gold: A Comparison of Cyber and Regular Consumer Fraud in The United States

### 2017-01-18 paper/paper.tex
added .gitignore for latex, and spelling
abstract

While cybercrime is a relatively recent human invention, fraudulent activity has a much longer history. This paper investigates the differences in fraud reporting and targeting among United States based consumers by comparing fraudulent activity mediated by the Internet, with traditional forms of fraud. We find that \ck{ADD QUICK AND EASY CONCLUSIONS HERE}.

@@

In the United States, a total of more that 25 million people are victims of frauds \ck{per year?} \cite{anderson2013consumer}. These deceptive scams are a major cause of economic damages to users, not to mention the added externalities of wasted time and stress. These illicit practices have been a part of the underground economy for a while. Initially individuals were tricked by scam calls, mail or in-person fraudsters, however, the rise of the Internet has provided fraudulent entities with a more streamlined exposure to the overall population. A  survey report from 2014 indicated that 47\% of Americans had been victims of an online identity theft. Another recent report released by the  Federal Trade Commission (FTC) revealed that debt collection, identity theft, and impostor scams contribute towards 56\% of the total frauds complaints in 2015 \cite{ftcpress2016} . With the number of Internet users on the rise, the number of cyber frauds is likely to increase over the next couple of years \cite{perharassment}. To deal with the increasing trend, the Federal Bureau of Investigation (FBI), alongside the FTC, has established a similar complaint portal known as the IC3, for the collection of specifically Internet-based fraudulent practices \cite{fbiic3}. This emerging trend of deceptive practices necessitates their study in order to evaluate and mitigate the harm caused to victims individuals.

@@

\begin{table*}[h]
\noindent
\centering
\begin{tabular}{K{0.18\linewidth}|p{0.72\linewidth}}
\hline
\multicolumn{1}{c}{\bfseries Data Field} & \multicolumn{1}{c}{\bfseries Field Description}
\\ 
\hline
\hline
Agency Name & The complaint collection agencies associated with the FTC.\\
\hline
Zip code Information & The zip code of the victim and the fraudulent entity.\\
%This provided us with low level granularity for analysis, however, the latter was only available for a subset of the complaints\\
\hline
Contact Method & The primary channel used by the fraudulent entity to contact the victim e.g. Internet, phone, mail.\\
\hline
Fraud Description & A description of nature the fraud, and its type e.g. credit card, fake product, debt collection.\\
\hline
Fraud \&
 Reporting Date & The dates when the fraud initially occurred and the date on which is was reported.
\\
\hline
\end{tabular}
\vspace{8pt}
\caption{The description of the data fields that were primarily used in the for data calibration and analysis}\label{ftcdata}
\vspace{-20pt}
\end{table*}

@@

\begin{table}[h]
\centering
\begin{tabular}{c|c}
\hline
{\bfseries Description} & \multicolumn{1}{c}{\bfseries Value}
\\
\hline
\hline
\% Cyber Complaints &52.1\\
\hline
\% Regular Complaints & 47.9\\
\hline
Moth with Most Compliants & July 2013\\
\hline
Moth with Least Compliants & Feb 2013\\
\hline
\end{tabular}
\vspace{8pt}
\caption{t-test results for cyber and regular frauds in different periods}\label{ttest}
\vspace{-10pt}
\end{table}

>>>>

While cybercrime is a relatively recent human invention, fraudulent activity has a much longer history. This paper investigates the differences in fraud reporting and targeting among United States based consumers by comparing fraudulent activity mediated by the Internet, with traditional forms of fraud. Both categories of frauds experience a decrease during the holiday season. We study Interestingly, individuals who are victimized via traditional methods are more likely to complain using phone or in-person, while 82\% of the cyber victims use an online resource to  generate fraud complaints.

 \ck{ADD QUICK AND EASY CONCLUSIONS HERE}.

@@

In the United States, a total of more that 25 million people are victims of frauds per year. \cite{anderson2013consumer}. These deceptive scams are a major cause of economic damages to users, not to mention the added externalities of wasted time and stress. These illicit practices have been a part of the underground economy for a while. Initially individuals were tricked by scam calls, mail or in-person fraudsters, however, the rise of the Internet has provided fraudulent entities with a more streamlined exposure to the overall population. A  survey report from 2014 indicated that 47\% of Americans had been victims of an online identity theft. Another recent report released by the  Federal Trade Commission (FTC) revealed that debt collection, identity theft, and impostor scams contribute towards 56\% of the total frauds complaints in 2015 \cite{ftcpress2016} . With the number of Internet users on the rise, the number of cyber frauds is likely to increase over the next couple of years \cite{perharassment}. To deal with the increasing trend, the Federal Bureau of Investigation (FBI), alongside the FTC, has established a similar complaint portal known as the IC3, for the collection of specifically Internet-based fraudulent practices \cite{fbiic3}. This emerging trend of deceptive practices necessitates their study in order to evaluate and mitigate the harm caused to victims individuals.

@@

\begin{table*}[h]
\noindent
\centering
\begin{tabular}{K{0.18\linewidth}|p{0.72\linewidth}}
\hline
\multicolumn{1}{c}{\bfseries Data Field} & \multicolumn{1}{c}{\bfseries Field Description}
\\ 
\hline
\hline
Agency Name & The complaint collection agencies associated with the FTC.\\
\hline
Zip code Information & The zip code of the victim and the fraudulent entity.\\
%This provided us with low level granularity for analysis, however, the latter was only available for a subset of the complaints\\
\hline
Contact Method & The primary channel used by the fraudulent entity to contact the victim e.g. Internet, phone, mail.\\
\hline
Fraud Description & A description of nature the fraud, and its type e.g. credit card, fake product, debt collection.\\
\hline
Fraud \&
 Reporting Date & The dates when the fraud initially occurred and the date on which is was reported.
\\
\hline
\end{tabular}
\vspace{8pt}
\caption{data field descriptions primarily used for data calibration and analysis}\label{ftcdata}
\vspace{-20pt}
\end{table*}

@@

\begin{table}[h]
\centering
\begin{tabular}{c|c}
\hline
{\bfseries Description} & \multicolumn{1}{c}{\bfseries Value}
\\
\hline
\hline
\% Cyber Complaints &52.1\\
\hline
\% Regular Complaints & 47.9\\
\hline
Month with Most Complaints & July 2013\\
\hline
Month with Least Complaints & Feb 2013\\
\hline
\end{tabular}
\vspace{8pt}
\caption{t-test results for cyber and regular frauds in different periods}\label{ttest}
\vspace{-10pt}
\end{table}
